Combining formal and distributional models of predicate semantics

### 2016-03-03 paper/paper.tex
writing background section
Former-commit-id: 5bb9d830736c482fd42cb272f37ddebbdfdf29c5
Former-commit-id: 0736fea577294a02f29dd7c4c39cf06135a1cc70

\documentclass[11pt]{article}
\usepackage{acl2016}
\usepackage{times}
\usepackage{latexsym}
\usepackage{color}

@@

@@

\mattnote{TODO - Jayant will write this}

>>>>

\documentclass[11pt]{article}
\usepackage{acl2016}
\usepackage{amsmath}
\usepackage{times}
\usepackage{latexsym}
\usepackage{color}

@@

\newcommand{\true}[0]{\predicate{true}}
\newcommand{\false}[0]{\predicate{false}}

@@

This system predicts a denotation $\gamma$, i.e., a set of Freebase
entities, for a natural language text $s$, using the following
probabilistic model:

\begin{align*}
P(\gamma | s) & = \sum_w \sum_\ell P(\gamma | \ell, w) P(w ; \theta, \phi) P(\ell | s)
\end{align*}

In this equation, $P(\ell | s)$ represents a semantic parser that
produces a logical form $\ell$ given the text $s$. This logical form
contains open-vocabulary predicates, as in the ``architect'' example
above. This factor is deterministic, meaning that it assigns
probability one to a single logical form. The second factor $P(w)$
represents a distribution over possible worlds $w$, where a world is
an assignment of truth values to all possible predicate
instances. This distribution can be efficiently represented using a
probabilistic database because each predicate instance is treated as
an independent random variable. The truth probabilities of these
variables are predicted using a matrix factorization approach similiar
to that of Riedel et al., \shortcite{riedel-2013-mf-universal-schema};
$\theta$ and $\phi$ represent the parameters of the matrix
factorization. Finally, $P(\gamma | \ell, w)$ represents the
deterministic evaluation of the logical form $\ell$ on the world $w$
to produce a denotation $\gamma$. This factor is similar to other
semantic parsing work, where it represents the deterministic
evaluation of a query against a given database. Algorithms from the
probabilistic databases community enable us to efficiently compute the
marginal probability that an entity $e$ is in the denotation of a
sentence, $\sum_\gamma P(e \in \gamma | s)$, for a large number of
queries; we refer the reader to (TODO).

The training problem in this model is to learn the parameters $\theta$
and $\phi$ of the probabilistic database. The probability of a
category instance $c(e)$ and relation instance $r(e_1, e_2)$ in the
database is given by:

\begin{align*}
  P(c(e) & = \true{}) & = & \sigma ( \theta_c^T \phi_e ) \\
  P(r(e_1, e_2) & = \true{}) & = & \sigma ( \theta_r^T \phi_{(e_1, e_2)} )
\end{align*}

The parameters $\theta$ represent $k$-dimensional embeddings of each
category and relation predicate, and the parameters $\phi$ represent
$k$-dimensional embeddings of each Freebase entity and entity
pair. The probability of a predicate instance is the inner product of
the corresponding predicate and entity embeddings. Thus, predicates
with nearby embeddings will have similar distributions over the
entities in their denotation.

The parameters $\theta$ and $\phi$ are learned from predicate/entity
co-occurrence counts collected from a large corpus. The training data
consists of query/entity pairs, $\{(\ell^i, e^i)\}$, where the entity
$e^i$ is observed to be an element of $\ell^i$'s denotation. These
pairs are collected by semantically parsing entity-linked sentences
and simplifying the resulting logical forms to identify conjunctions
of predicates where all arguments are bound to specific
entities. Extracting an entity from this conjunction results in a
query/entity pair. The model is trained using a ranking objective that
learns to rank the observed entities for each query above unobserved
entities.
